Supporting Information:\\Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-12-11 Manuscript/SI.tex
SI cleanup

\begin{table}[!h]
  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si}
  {\color{red}Joe, please deliver the citations for all experimental area per molecule values or remove them.}}
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids &        &  \\
    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
    ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
    ~ TIP3p          & 66.2$\pm$ 0.6   &  313       \\
    ~ TIP3p-FB       & 64.8$\pm$ 0.6   &  313       \\
    ~ TIP4p-FB       & 65.6$\pm$ 0.6   &  313       \\
    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
    %oMM small patch           & 63.65  &         \\
    %oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment   & 62.7  &  293    \\
    experiment \cite{kucerka11} & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    %experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular} \\
\end{table}

@@

The results from simulations are shown in Fig.~\ref{fig:cacl-bind}
together with the line fitted to experimental data by Altenbach and Seelig~\cite{altenbach84}.
Both results are in agreement with the prediction of the ternary complex model.
The small discrepancy between the results from experiments and simulations probably
arise from difference in the evaluation of the concentrations and
inaccuracy of Poisson-Boltzmann theory for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
In conclusion, the results suggest that the almost equal probabilities for
\ce{Ca^{2+}} to form complexes with one to three lipids detailed in the main text is in line with an averaged interpretation of the experimental observations which supported the ternary complex model.
\newpage
\begin{figure}[!h]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Fits of experimental~\cite{altenbach84} and ECC-POPC simulation data to
    the prediction of the ternary complex model.
    {\color{red} Joe, please change the axes labels to correspond the notation in the text, i.e. $C_b$ and $C_I$,
    and write in the title of simulation fit (red dashed line) that it is the fit to simulation.}
    %Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
    %that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    %provides a good fit to experimental measurements~\cite{altenbach84}
    %and it also provides a good fit to our simulation data. 
    %This supports our statement that the success of the ternary complex model in the experiments
    %can be understood with the atomistic detail of our simulations as an average 
    %stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
    %Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    %and, hence, the observed slope of the linear relationship is slightly different.
    }
\end{figure}

@@

\begin{figure}[!p]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_surf_GMX-oMM_compar.pdf}
  \caption{\label{fig:ordPars_surf_GMX_oMM_compar}
    Changes of headgroup and glycerol carbon $g_3$ order parameters, and P-N vector orientation as a function of
    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
    from simulations with the ECC-lipids models
    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
    compared with the experimental values from \cite{scherer89}.
    The size of the markers for the head group order parameters correspond to
    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
    while the error estimate for simulations is $\pm 0.005$.
  }
\end{figure}

>>>>

\begin{table}[!h]
  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-POPC &        &  \\
    ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
    ~ TIP3p          & 66.2$\pm$ 0.6   &  313       \\
    ~ TIP3p-FB       & 64.8$\pm$ 0.6   &  313       \\
    ~ TIP4p-FB       & 65.6$\pm$ 0.6   &  313       \\
    \hline
    experiment \cite{kucerka11}  & 62.7$\pm$ 1.3  &  293    \\
                                 & 64.3$\pm$ 1.3  &  303    \\
                                 & 67.3$\pm$ 1.3  &  323    \\
                                 & 68.1$\pm$ 1.4  &  333    \\
    %experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular} \\
\end{table}

@@

The results from simulations are shown in Fig.~\ref{fig:cacl-bind}
together with the line fitted to experimental data by Altenbach and Seelig~\cite{altenbach84}.
Both results are in agreement with the prediction of the ternary complex model.
The small discrepancy between the results from experiments and simulations probably
arise from difference in the evaluation of the concentrations and
inaccuracy of Poisson-Boltzmann theory for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
In conclusion, the results suggest that the almost equal probabilities for
\ce{Ca^{2+}} to form complexes with one to three lipids detailed in the main text is in line with an averaged interpretation of the experimental observations which supported the ternary complex model.
\newpage
\begin{figure}[!h]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Fits of experimental~\cite{altenbach84} and ECC-POPC simulation data to
    the prediction of the ternary complex model.
    %{\color{red} Joe, please change the axes labels to correspond the notation in the text, i.e. $C_b$ and $C_I$,
    %and write in the title of simulation fit (red dashed line) that it is the fit to simulation.}
    %Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
    %that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    %provides a good fit to experimental measurements~\cite{altenbach84}
    %and it also provides a good fit to our simulation data. 
    %This supports our statement that the success of the ternary complex model in the experiments
    %can be understood with the atomistic detail of our simulations as an average 
    %stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
    %Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    %and, hence, the observed slope of the linear relationship is slightly different.
    }
\end{figure}

@@

\begin{figure}[!p]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_surf_GMX-oMM_compar.pdf}
  \caption{\label{fig:ordPars_surf_GMX_oMM_compar}
    Changes of headgroup and glycerol carbon $g_3$ order parameters, and P-N vector orientation as a function of
    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
    from simulations with the ECC-POPC model
    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
    compared with the experimental values from \cite{scherer89}.
    The size of the markers for the head group order parameters correspond to
    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
    while the error estimate for simulations is $\pm 0.005$.
  }
\end{figure}
